---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---


```{r header, echo=FALSE,results = "hide"}
library(ggplot2)
library(dplyr)
library(lubridate)
```

#Loading and preprocessing the data

```{r read}
SrcFilename<-"repdata-data-activity.zip"
if (!file.exists(SrcFilename)){
  #  quit(save = "ask", status = 0,
  #       runLast = TRUE)
  print (paste("file ",SrcFilename,"File not found"))
}
unzip(SrcFilename)
ActivityData<-read.csv("activity.csv",header=TRUE,sep=",")
```

#What is mean total number of steps taken per day

```{r aggregatedata}
SumDataActivity<-aggregate(steps~date,ActivityData,sum)
```

1.) Plot the histogram of the total number of steps taken each day.

```{r makehistogram}
p<- ggplot(SumDataActivity,aes(x = steps))+ 
  geom_histogram(fill="blue") +
  labs(x="Total number of steps per day", y = "Count")
print(p)
```

2.) Mean and median of the number of steps
``` {r meanmedian}
MeanVal<-mean(SumDataActivity[,2])
MedianVal<-median(SumDataActivity[,2])
print(paste("Mean: ",MeanVal))
print(paste("Median: ",MedianVal))
```

#What is the average daily activity pattern?

```{r averageDailyPat }
AvgDailyIntervalPat<-aggregate(steps~interval,ActivityData,mean)
with(AvgDailyIntervalPat,{ 
  plot(interval,steps,type="l",xlab="time [sec]",ylab="Avg number of steps per day",lwd="2",main="Average activity",col="red") 
})
```

- Get the interval with max number of steps over all days

```{r}
indx<-which.max(AvgDailyIntervalPat[,2])
Intervalbegin<-AvgDailyIntervalPat[indx,1]
Intervalend<-Intervalbegin+5
print(paste("Interval of max mean steps:",Intervalbegin,"-",Intervalend))
```

#Imputing missing values
- Count Number of rows with NA

```{r}
NumCompCases<-sum(complete.cases(ActivityData))
NumNas<-sum(!complete.cases(ActivityData))
print(paste("Number of rows with NAs:",NumNas))
```

- Fill in the NAs
The strategy that I will use is to take replace the NAs by the mean for that day.
The new data set is called ActivityFullData.

```{r FillData}
MeanByDay<-aggregate(steps~date,ActivityData,mean)
ActivityFullData<-transform(ActivityData,steps=ifelse(is.na(steps),
              MeanByDay[as.Date(MeanByDay$date)%in%as.Date(ActivityData$date),2],steps))
```

- Make a new histogram  of the total number of steps taken each day
```{r makehistogram1}
SumDataActivityFull<-aggregate(steps~date,ActivityData,sum)
p1<- ggplot(SumDataActivityFull,aes(x = steps))+ 
  geom_histogram(fill="green") +
  labs(x="Total number of steps per day", y = "Count")
print(p1)
```

-New Mean and median of the number of steps

``` {r meanmedianNew}
MeanValNew<-mean(SumDataActivityFull[,2])
MedianValNew<-median(SumDataActivityFull[,2])
print(paste("New mean:",MeanValNew))
print(paste("New median:",MedianValNew))
```

The mean and median values are not affected by imputing the NAs with 
means over that corresponding day.

#Are there differences in activity patterns between weekdays and weekends?
- First create the grouped data and the factors weekends and weekdays

```{r}
ActivityFullData<-mutate(ActivityFullData,weekdays(as.Date(date)))
names(ActivityFullData)[4]<-c("DayType")
ActivityFullData<-transform(ActivityFullData,DayType=ifelse(DayType %in% c("Saturday","Sunday"),"weekend","weekday"))
ActivityFullData$DayType<-as.factor(ActivityFullData$DayType)
```

#Are there differences in activity patterns between weekdays and weekends?

```{r}
AvgFullData<-aggregate(steps~interval+DayType,ActivityFullData,mean)
p3<-ggplot(AvgFullData,aes(interval,steps,color=DayType)) +
  geom_line() +
  labs(x="Interval", y = "Number of steps") +
  facet_wrap(~DayType,nrow=2)
print(p3)
```

-There is in general a higher activity on the weekends compared to weekdays.
